A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-08-29 paper/paper.tex
paper: MAJ

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommandx{\myframe}[2][1,2]{\prim{frame}[][#1][#2][#2][]}
\newcommandx{\mylayer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\myfiller}[2][1,2]{\prim{filler}[][#1][#2][][]}

@@

\subsection{Iterated parametricity}
A full 2-cube is made of two full 1-cubes connected by a lifted 1-cube made of two lines connecting the former points and a square connecting the former lines:
\begin{equation*}
  \begin{array}{l}
    \begin{array}{lll}
      \mbox{\textit{squares}}     \\
      [3mm]\mbox{\textit{lines}}  \\
      [3mm]\mbox{\textit{points}} \\
      [1mm]\end{array}
    \framebox{$\begin{array}{lll}
                     \overbrace{\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[1][0]$}} & \framebox{$a_{0\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{00}$}~\framebox{$a_{01}$}}   &
                  \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[1][0]\!\!\!\!\!\!\!\!$}}
                  \\
                \end{array}
                         $}}~
                     {\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[1][0]$}} & \framebox{$a_{1\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{10}$}~\framebox{$a_{11}$}}   &
                  \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[1][0]\!\!\!\!\!\!\!\!$}}
                \end{array}
                         $}}}}^{\mylayer[2][0]}~\stackrel{\raisebox{9.5mm}{$\framebox{$
                               \begin{array}{ll}
                      \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[2][1]$}}                   & \framebox{$a_{\unitpoint\unitpoint}$} \\
                      [1mm]\overbrace{\framebox{$a_{\unitpoint 0}$}~\framebox{$a_{\unitpoint 1}$}} &
                      \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[2][1]\!\!\!\!\!\!\!\!$}}
                    \end{array}$}$}}
                     {\raisebox{-0.4cm}{{\scriptsize \hspace{2.5cm}$\myfiller[2][0]\!\!\!\!\!\!\!$}}}
                   \end{array}$}
    ~~~\parbox{3.5cm}{{\it additionally, each atomic component at dimension $n$ is a $\myfiller[n][n]$}} \\
    [-1mm]\qquad\qquad~~
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize $\myframe[1][1]$}
    \quad~~
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet \\
    \end{tikzcd}
    \mbox{\normalsize $\myframe[1][1]$}                                                                  \\
    [-2mm]\qquad\qquad
    \begin{tikzcd}
     \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize$\myframe[2][1]$}                                                                   \\
    [-2mm]\qquad\qquad
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize$\myframe[2][2]$}                                                                   \\
  \end{array}
\end{equation*}

@@

\begin{align*}
  \mathsf{eq\_refl}        & : \forall x, x = x                                           \\
  \mathsf{UIP\_refl\_unit} & : \forall x : \unitpoint = \unitpoint, x = \mathsf{eq\_refl}
\end{align*}

>>>>

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommandx{\myframe}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\mylayer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\myfiller}[2][1,2]{\prim{filler}[][#1][#2][][]}

@@

\subsection{Iterated parametricity}
One step of parametricity can be seen as interpreting a point by a line. Iterating parametricity shall lead to map terms typed in a context with $n$ declarations of an axis as $n$-cubes. We shall interpret a
typing derivation with $n$ axis declarations as an $n$-cube of typing derivations in the target. For instance, a typing derivation with exactly one declaration of an axis will be interpreted by 3 typing derivations in the target, two of them typing points and the third one typing a proof, seen as a line, together with the fact that these two points are related (as in ordinary parametricity), while a typing derivation with $n$ declarations of an axis shall be interpreted as a $3^n$ derivations, split into $2^n$ derivations of points.

A full 2-cube is made of two full 1-cubes connected by a lifted 1-cube made of two lines connecting the former points and a square connecting the former lines:
\begin{equation*}
  \begin{array}{l}
    \begin{array}{lll}
      \mbox{\textit{squares}}     \\
      [3mm]\mbox{\textit{lines}}  \\
      [3mm]\mbox{\textit{points}} \\
      [1mm]\end{array}
    \framebox{$\begin{array}{lll}
                     \overbrace{\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[1][0]$}} & \framebox{$a_{0\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{00}$}~\framebox{$a_{01}$}}   &
                  \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[1][0]\!\!\!\!\!\!\!\!$}}
                  \\
                \end{array}
                         $}}~
                     {\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[1][0]$}} & \framebox{$a_{1\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{10}$}~\framebox{$a_{11}$}}   &
                  \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[1][0]\!\!\!\!\!\!\!\!$}}
                \end{array}
                         $}}}}^{\mylayer[2][0]}~\stackrel{\raisebox{9.5mm}{$\framebox{$
                               \begin{array}{ll}
                      \raisebox{0cm}{\hspace{.5cm}{\scriptsize$\mylayer[2][1]$}}                   & \framebox{$a_{\unitpoint\unitpoint}$} \\
                      [1mm]\overbrace{\framebox{$a_{\unitpoint 0}$}~\framebox{$a_{\unitpoint 1}$}} &
                      \raisebox{-0.1cm}{{\scriptsize \hspace{0.1cm}$\myfiller[2][1]\!\!\!\!\!\!\!\!$}}
                    \end{array}$}$}}
                     {\raisebox{-0.4cm}{{\scriptsize \hspace{2.5cm}$\myfiller[2][0]\!\!\!\!\!\!\!$}}}
                   \end{array}$}
    ~~~\parbox{3.5cm}{{\it additionally, each atomic component at dimension $n$ is a $\myfiller[n][n]$}} \\
    [-1mm]\qquad\qquad~~
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize $\myframe[1][1]$}
    \quad~~
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet \\
    \end{tikzcd}
    \mbox{\normalsize $\myframe[1][1]$}                                                                  \\
    [-2mm]\qquad\qquad
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize$\myframe[2][1]$}                                                                   \\
    [-2mm]\qquad\qquad
    \begin{tikzcd}
      \bullet\arrow[r, leftrightarrow] & \bullet\\
    \end{tikzcd}
    \mbox{\normalsize$\myframe[2][2]$}                                                                   \\
  \end{array}
\end{equation*}

@@

\begin{align*}
  \mathsf{eq\_refl}        & : \forall x, x = x                                                    \\
  \mathsf{UIP\_refl\_unit} & : \forall x : \unitpoint = \unitpoint, x = \mathsf{eq\_refl}          \\
  \mathsf{UIP\_refl\_bool} & : \forall b : \mathsf{bool}, \forall x : b = b, x = \mathsf{eq\_refl} \\
\end{align*}
